Analyzing Security-Cost Trade-off for a \\ Fully Homomorphic Encryption Scheme

### 2013-06-16 Report/article.tex
alpha version article.tex

\usepackage{listings}
\usepackage{algorithm2e}
\usepackage{verbatim}
\begin{document}
\title{Analysis of Security-Cost Trade-off of Fully Homomorphic Encryption Schemes}

%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

@@

\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.
\maketitle
\begin{abstract}
This paper investigates the feasibility of transformation of a (possibly) non-straight-line program (on unencrypted data) to a straight-line program on encrypted data. We present an analysis of security-cost trade-off for homomorphic encryption schemes on such programs. Analysis is based on the measurements (CPU time and  Wall time) taken for programs : performing \texttt{XOR} of an $n$-bit sequence, determining the majority bit of an $n$-bit sequence, evaluating the sum of $n$-bit integers, \textcolor{orange}{sum of integers with any number of bits} and sorting $n$ bit sequences of length $nbits$ (requires verifying the validity of a boolean predicate). The evaluation is performed on an available implementation ``Scarab library'' of a fully homomorphic encryption scheme.

@@

The notion of \textit{fully homomorphic encryption scheme}, orginally called a \textit{privacy homomorphism} was introduced by Rivest et al. \cite{rivest78} shortly after the invention of the RSA by Rivest, Shamir and Adleman \cite{Rivest78amethod}.  Basic RSA is a multiplicative homomorphic encryption scheme -- i.e. given the public parameters $pk:=(e, N)$ and two messages $m_0, m_1$, their encryption $c_0, c_1$ verify $c_0c_1=(m_0m_1)^e \ \textrm {mod}\ N$. Hence, without knowing the associated secret key, an external agent can compute the encryption of the product of the messages.

Fully homomorphic encryption scheme extends the above property to addition. A Fully homomorphic scheme is a tuple of an ecryption $\mathcal{E}$, decryption $\mathcal{D}$ algorithm together with an efficient algorithm $Eval_\mathcal{E}$ that evaluates any circuit $C$ on any ciphertexts $c_i \leftarrow \mathcal{E}(pk, m_i)$. The property that it can evaluate the encryption of addition and multiplication of two plaintexts (without access to the secret key) allows it to arbitrarily compute any function on encrypted data without the decryption key. However prior to \cite{homenc}, we did not have a viable construction.

@@

The immediate question that is raised, is how do we perform the transformation and how does the circuit size change to adapt to computations on encrypted data. We provide answer to this question by analyzing circuit complexity for finding the minimum-maximum of two $n$-bit sequences in \autoref{sec:ex} and for some other opertation in \autoref{sec:bm}. Moreover, the cost of evaluation depends on the depth of the circuit representing the $Eval_{\mathcal{E}}$ function. Some researchers have improved and proposed other schemes such as \cite{cryptoeprint:2009:571} and \cite{cryptoeprint:2011:277} which decrease the circuit complexity. In \autoref{Sec:Eval} we expermentally evaluate an implementation of  \cite{cryptoeprint:2009:571} and analyze the cost-security trade-off for $Eval_{\mathcal{E}}$ varying from \texttt{XOR} of the bits of an $n$ bit integer to sum of bounded and unbounded integers to sorting of $n$ integers. Three different sorting algorithms are tested.  
Through these experiments, we also provide information on the additional cost to be paid (for these computations) to increase the security of the cyptosystem.

@@

The homomorphic encryption scheme proposed by Gentry \cite{homenc} is based on ideal lattices. The Gentry's non-determi-nistic scheme builds on a \textit{somewhat homomorphic} scheme which can perform additions and multiplications on encrypted data till a bounded depth. The non-determinism introduced in the form of a noise renders the scheme secure. But, the noise increases with the depth  of the evaluation circuit. To counter this problem, Gentry introduces \textit{bootstrapping} technique which ensures unbounded depth operatability of the scheme. Smart Vercauteren \cite{cryptoeprint:2009:571} have proposed an integer based approach and we use one of its implementation detailed in \cite{perl:poster} for our experiments.

@@

The scheme proposed in \cite{cryptoeprint:2009:571} has 3 parameters:($N, \nu, \mu$ ). The somewhat homomorphic scheme uses 5 algorithms: (\texttt{KeyGen}, \texttt{Encrypt}, \texttt{Decrypt}, \texttt{Add}, \texttt{Mult}). Algorithm \autoref{Code:SV} presents the \texttt{KeyGen} algorithm.

@@

\texttt{Encrypt}($m \in \{0,1\} , pk$): 
\newline \phantom{x}\hspace{3ex}  $R(x)=_{R}(B_{\infty , N}(\mu/2)$
 \; $C(x)=m+2\cdot R(x)$ 
\newline \phantom{x}\hspace{3ex} return  $c=C(\alpha)$ mod $p$
\newline \texttt{Decrypt}($c, sk$):
\newline \phantom{x}\hspace{3ex} $m = (c - \lfloor c \cdot B/p \rceil )$ mod $2$;
\phantom{x}\hspace{1ex} return $m$
\newline \texttt{Add}($c_1$, $c_2$, $pk$):
\newline \phantom{x}\hspace{3ex} $c_3=c_1+c_2$ mod $p$
;  \phantom{x}\hspace{1ex}  return $c_3$
;
\newline \texttt{Mult}($c_1$, $c_2$, $pk$):
 \newline \phantom{x}\hspace{3ex}  $c_3=c_1.c_2$ mod $p$
 ; \phantom{x}\hspace{1ex}  return $c_3$
;

\textbf{Security of the scheme} The scheme is semantically secure under the assumption that the \textit{Polyomial Coset Problem} is hard. One wayness of the encryption is believed to be true under the assumption that the \textit{Bounded distance decoding problem} BDDP is hard while key recovery is assumed to be difficult under the assumption that the \textit{Small Principal Ideal Problem} is hard. Solutions to all of these problems have either a sub-exponential algorithm or in general take exponential time.

@@

\textbf{SWHE to FHE :} We recall that the authors have proved that the multiplicative depth $d$ of SWHE  verifies \[d \cdot log(2) \leq log( log(\frac{\nu}{2\cdot \sqrt{N}}) ) - log(log( N\mu))\]
when $F(x)= x^N+1$. Turning the SWHE to FHE is achieved by recasting Gentr's method for the parameters of this scheme. To achive this the authors propose to generate $s_1$ uniformly random integers $B_i$ in $[-p, \ldots, p]$ such that there exists a subset $S$ of $s_2$ elements with $\sum_{j \in S}{B_j}= B$ over integers. By defining $sk_i =1$ iff $i\in S$ and taking $c_i=Encrypt(sk_i,pk)$ we generate a new public key which contains hint of the secret key $sk$. The public key now comprises of \[(p, \alpha, s_1, s_2, \{c_i, B_i\}_{i=1}^{s1})\].

To ensure that the hint doesn't provide considerable advantage to an attacker, the underlying \textit{sparse subset-sum problem} SSSP must remain remain hard. If we take $s_1$ to be slightly greater than $p$, then we need to select $s_2$ such that $\sqrt {{s_1}  \choose {s_2} } \leq 2 ^{N/\lambda}$ so as to ensure that SSSP difficulty is at least as difficult as the difficulty of the BDDP underlying the SWHE scheme.
\subsection{Considered Implementations}
We used the open source implementation Scarab library \cite{hcrypt}  version $1.0.0$ of the Smart-Vercauteren scheme and is a part of \textsc{hcryptProject}. The implementation is in beta version. The software requires GMP \cite{gmp}, The GNU Multiple Precision Arithmetic Library; FLINT \cite{flint} , Fast Library for Number Theory; MPIR \cite{mpir}, Multiple Precision Integers and Rationals and MPFR \cite{mpfr}, multiple-precision floating-point computations with correct rounding. Linux is supposed to be the only supported platform. The libraries are in implemented in  C and the installation guide is provided with the package. \autoref{funs} presents the list of functions provided by the implementation with their description. The library provides a file \texttt{test.c} where a user can write his own function to test different operations.
\begin{table*}
\caption{Function available in Scarab}
\begin{tabular}{|l|l||}
  \hline
  \textbf{function prototype} & \textbf{semantics}  \\
  \hline
 \texttt{fhe\_keygen(fhe\_pk\_t pk, fhe\_sk\_t sk);}  & 	Generate a keypair \\
\texttt{fhe\_encrypt(mpz\_t c, fhe\_pk\_t pk, int m);} &	Encrypt a message (0 or 1) \\
\texttt{fhe\_decrypt(mpz\_t c, fhe\_sk\_t sk);} &	Decrypt a cyphertext\\
\texttt{fhe\_recrypt(mpz\_t c, fhe\_pk\_t pk, fhe\_sk\_t sk); }	&Recrypt a cyphertext ( ``refreshing'' it) \\
\texttt{fhe\_add(mpz\_t res, mpz\_t a, mpz\_t b, fhe\_pk\_t pk);} &	Add cyphertexts (= XOR) \\
\texttt{fhe\_mul(mpz\_t res, mpz\_t a, mpz\_t b, fhe\_pk\_t pk);} &	Multiply cyphertexts (= AND) \\
\texttt{fhe\_fulladd(mpz\_t sum, mpz\_t c\_out, mpz\_t a, mpz\_t b,} &	Add with carry in and carry out \\
\phantom{x}\hspace{12ex} \texttt{ mpz\_t c\_in, fhe\_pk\_t pk);} & \\
\texttt{fhe\_halfadd(mpz\_t sum, mpz\_t c\_out, mpz\_t a,} &	Add with carry out \\
\phantom{x}\hspace{12ex} \texttt{ mpz\_t b, fhe\_pk\_t pk);} & \\  
\hline
\end{tabular}
\label{funs}
\end{table*}

To simplify out computational circuit added, we added two basic functions \texttt{or(mpz\_t res, mpz\_t a, mpz\_t b, fhe\_pk\_t pk);} and \texttt{not(mpz\_t res, mpz\_t a, fhe\_pk\_t pk);} which compute the logical or of encrypted bits and the negation of an encrypted bit respectively.

We time the various algorithms on two desktop machines. Both these machines were x86-64 platform running linux kernel $3.2.0$-$45$-generic  and used GCC $4.6.3$ C compiler.  The first machine housed Intel Core i3 M330 (2.13GHz x 4) processor and will be referred in this article as machine M1 while the other used  housed Intel Core i5 2430 M(2.40 x4) processor and will be referred as machine M2.

\subsection{Intial set-up cost}
Here we mesure the performance in terms of run time for KeyGen, Encrypt, Add, Mult, Recrypt, Decrypt.\\
\begin{table}[htb]
  \centering
  \caption{Run time (ms)}
\resizebox{9 cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|c|}
  \hline
  \multicolumn{8}{|c|}{$\mu = 4$ , $\eta = 2^{384}$} \\
  \hline
  $N=2^{n}$ & $log_{2}(p)$ & KeyGen & Encrypt & Add & Mult & Recrypt & Decrypt \\
  \hline
  $2^{1}$ & 764 & 99 & 0.057 & 0.0 & 2.372 & 2.371 & 0.0 \\
  $2^{2}$ & 1535 & 221 & 0.072 & 0.0 & 	3.012 & 3.023 & 0.001\\
  $2^{3}$ & 3077 & 1131 & 0.079 & 0.001 & 3.396 & 3.380 & 0.004 \\
  $2^{4}$ & 6155 & 8560 & 0.135 & 0.002 & 5.881 & 5.889 & 0.012\\
  $2^{5}$ & 12329 & 852849 & 0.477 & 0.005 & 20.230 & 20.057 & 0.028\\
  \hline
\end{tabular}
\normalsize
\label{common cost}
}
\end{table}

@@

\textbf{Encryption and Decryption}\\

\section{Circuit Transformation}
As previously discussed, a user sends his program which works on unencrypted data and his encrypted input to the cloud. The cloud then performs the demanded computation and returns the result. However, the program in the same form cannot always be used for computations on encrypted data.

Consider a program involves the following test on two inputs : \texttt{if(a > b) instruction\_1; else instruction\_2;}. As the cloud does not know the result of the test\texttt{ a' > b'} where \texttt{a'} and \texttt{b'} are the encryptions of \texttt{a} and \texttt{b}, he does not know which of the instructions he has perform. In fact he performs both the instrunctions by transforming the actual program into : \texttt{test(a'> b')*} \texttt{instruction\_1 +}$\neg$\texttt{test(a'> b')*instruction\_2}, where \texttt{test()}  is a function on encrypted inputs \texttt{a'}, \texttt{b'} and returns an encryption of $1$ iff \texttt{a > b}. Hence, if the program used has braching i.e. it involves \texttt{if(cond\_0)} \texttt{\{inst\_0\} else if(cond\_1)\{inst\_1\}}\linebreak \texttt{ ...else if(cond\_n-2) \{inst\_n-2\}} \texttt{else\{inst\_n-1 \}}; the cloud has to perform all the $n$ instructions i.e. the size of the circuit changes from a constant size to a size linear in the number of tests.

@@

\textbf{Another transformation}: We consider the following program that a cloud receives : \\
 \phantom{x} \texttt{i=0, res=0;}
\newline \phantom{x} \texttt{do\{ }
\newline \phantom{x} \hspace{9ex} \texttt{i++;} 
\newline   \phantom{x} \} \texttt{while(i < n \&\& cond\_i);} 
\newline  \phantom{x} \texttt{res = f(i)};\\
In this case, the cloud needs to transform a loop with condititonal stopping into a loop with unconditional branching. We propose the following possible transformation : \\
 \phantom{x} \texttt{endLoop=0', res=0', i=0;}
\newline \phantom{x} \texttt{do\{ }
\newline \phantom{x} \hspace{6ex} \texttt{endLoop = $\neg$cond\_i;}
\newline \phantom{x} \hspace{6ex} \texttt{res = $\neg$endLoop*f(i)+($\neg$cond\_i)*res;}
\newline \phantom{x} \hspace{6ex} \texttt{i++;} 
\newline   \phantom{x} \} \texttt{while(i < n);}

\textbf{Correctness of the code :} If \texttt{cond\_i} is always \texttt{true}, \texttt{endLoop} is always false hence the \texttt{res=f(n)}. If at the $t^{th}$iteration, \texttt{cond\_i} is \texttt{false}, \texttt{endLoop} becomes \texttt{false} and hence \texttt{res} contains its previous value $f(t-1)$.

@@

\subsection{ Didactic example :  Min-Max}
\label{sec:ex}

We suppose that a user wants to find the maximum and the minimum of two $n$-bit sequences $a = (a_0,a_1,\ldots, a_{n-1})$, $b=(b_0,b_1,\ldots, b_{n-1})$. He provides the following algorithm to the cloud :
\newline \phantom{x} \texttt{aIsGreater=0, i=0;}
\newline \phantom{x} \texttt{do\{ }
\newline \phantom{x} \hspace{6ex} \texttt{ if(a[i]==1 \&\& b[i]==0) 
\newline \phantom{x} \hspace{14ex}aIsGreater = 1;}
\newline \phantom{x} \hspace{7ex} \texttt{i++;} 
\newline   \phantom{x} \} \texttt{while(i < n);}

@@

 \caption{Min-Max on ciphertext \label{Code:algo}}

@@

Consider $x_i = a_ib_i+\neg(a_i)\neg(b_i) $. $x_i$ is $1$ iff $a_i$ and $b_i$ are equal. We note that, $bitEqual$ is the product of $x_i$ and is $1$ at the $t^{th}$ iteration iff $a_i=b_i$ for $i < t$. The term $\neg(b_i)a_i$ is $1$ iff $a_i=1$ and $b_i =0$. Hence at the $t^{th}$ iteration, the algorithm using $bitEqual$ tests if the previous bits were equal. Hence, $aIsGreater$ is $1$ iff at any stage  $a_t\neg(b_t)=1$ and the previous bits were equal.

Max and Min regenerate $a_i$ and $b_i$ and hence contain the correct result.  
\end{proof}
\textbf{Theoretical Cost:} Algorithm \autoref{Code:algo} uses $n$ \texttt{AND} gates, $n$ \texttt{OR} gates and $n$ \texttt{NOT} gates to find the larger of the two bit sequences and then to regenerate the maximum and the minuimum $4n$ \texttt{AND} gates and $2n$ \texttt{NOT} and \texttt{OR} gates. We note that reconstruction of Max and  Min is not required when working on non-encrypted data. Hence, this is an additional cost to pay to use the transformed program operating on encrypted data.

\textbf{Experimental Analysis :} We measure the run time from $n=1$ to $32$ bits, we find a linear cost which confirms the $\theta(n)$ complexity. During these measures the result of Min-Max is correct.
\begin{comment} 
\begin{figure}[!h] %on ouvre l'environnement figure
\centering
\includegraphics[width=5cm]{f2.eps} 
\caption{Run time(ms)} 
\label{image_f2} %l'étiquette pour faire référence à cette image
\end{figure}
\end{comment}
The cost of this algorithm on ciphertext is linear with bit number, like  the algorithm on plaintext, the extra cost in this case is encryption-decryption cost and especially the code generation which is very inpreductible and take often a long time in the SV-Scheme.

@@

We propose a set of benchmark functions to evaluate the implementations. These functions can be categorized into three classes : functions operating on bits, \textcolor{orange}{functions operating directly on integers or block of bits} and functions where branching is usually required i.e \texttt{if-then-else} condition is evaluated.

>>>>

\usepackage{listings}
\usepackage{algorithm2e}
\usepackage{verbatim}
\begin{document}
\title{Analyzing Security-Cost Trade-off for a \\ Fully Homomorphic Encryption Scheme}

@@

\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.
\maketitle
\begin{abstract}
This paper investigates the feasibility of transformation of a (possibly) non-straight-line program (on unencrypted data) into a straight-line program on encrypted data. We present an analysis of security-cost trade-off for a homomorphic encryption scheme on such programs. Analysis is based on the measurements (CPU time and  Wall time) taken for programs : performing \texttt{XOR} of an $n$-bit sequence,  evaluating the sum of two $n$-bit sequences, determining the majority bit of an $n$-bit sequence, finding the product of two $n \times n$ matrices, and sorting $n$ bit sequences of length $nbits$ (usually requiring validity of a boolean predicate). The evaluation is performed on an available implementation ``Scarab library'' of a fully homomorphic encryption scheme.

@@

The notion of \textit{fully homomorphic encryption scheme}, originally called a \textit{privacy homomorphism} was introduced by Rivest et al. \cite{rivest78} shortly after the invention of the RSA by Rivest, Shamir and Adleman \cite{Rivest78amethod}.  Basic RSA is a multiplicative homomorphic encryption scheme -- i.e. given the public parameters $pk:=(e, N)$ and two messages $m_0, m_1$, their encryption $c_0, c_1$ verify $c_0c_1=(m_0m_1)^e \ \textrm {mod}\ N$. Hence, without knowing the associated secret key, an external agent can compute the encryption of the product of the messages.

Fully homomorphic encryption scheme extends the above property to addition. A Fully homomorphic scheme is a tuple of an encryption $\mathcal{E}$, decryption $\mathcal{D}$ algorithm together with an efficient algorithm $Eval_\mathcal{E}$ that evaluates any circuit $C$ on any cipher-texts $c_i \leftarrow \mathcal{E}(pk, m_i)$. The property that it can evaluate the encryption of addition and multiplication of two plain-texts (without access to the secret key) allows it to arbitrarily compute any function on encrypted data without the decryption key. However prior to \cite{homenc}, we did not have a viable construction.

@@

The immediate question that is raised, is how do we perform the transformation and how does the circuit size change to adapt to computations on encrypted data. We provide answer to this question by analyzing circuit complexity for finding the minimum-maximum of two $n$-bit sequences in \autoref{sec:ex} and for some other operations in \autoref{sec:bm}. Moreover, the cost of evaluation increases with the depth of the circuit representing the $Eval_{\mathcal{E}}$ function. Some researchers have improved and proposed other schemes such as \cite{cryptoeprint:2009:571} and \cite{cryptoeprint:2011:277} which decrease the circuit complexity. In \autoref{Sec:Eval} we experimentally evaluate an implementation of  \cite{cryptoeprint:2009:571} and analyze the cost-security trade-off for $Eval_{\mathcal{E}}$ varying from \texttt{XOR} of the bits of an $n$ bit integer to sum of two $n$ bit sequences to sorting of $n$ sequences of $nbits$ each. Three different sorting algorithms are tested.  
Through these experiments, we also provide information on the additional cost to be paid (for these computations) when security of the cryptosystem is increased.

@@

The homomorphic encryption scheme proposed by Gentry \cite{homenc} is based on ideal lattices. The Gentry's non-deterministic scheme builds on a \textit{somewhat homomorphic encryption}(SWHE) scheme which can perform additions and multiplications on encrypted data till a bounded depth. The non-determinism introduced in the form of a noise renders the scheme secure. But, the noise increases with the depth  of the evaluation circuit and hence the decryption algorithm fails to produce the correct plain-text after a certain depth. To counter this problem, Gentry introduces \textit{bootstrapping} technique which ensures unbounded depth operatability of the scheme. Smart Vercauteren \cite{cryptoeprint:2009:571} have proposed an integer based approach and we use one of its implementations detailed in \cite{perl:poster} for our experiments.

@@

The scheme proposed in \cite{cryptoeprint:2009:571} has 3 parameters :($N, \nu, \mu$ ). The somewhat homomorphic scheme uses 5 algorithms: (\texttt{KeyGen}, \texttt{Encrypt}, \texttt{Decrypt}, \texttt{Add}, \texttt{Mult}). Algorithm \autoref{Code:SV} presents the \texttt{KeyGen} algorithm. This scheme will be referred to as SV scheme in this paper.

@@

\texttt{Encrypt}($m \in \{0,1\} , pk$): 
\newline \phantom{x}\hspace{3ex}  $R(x)=_{R}(B_{\infty , N}(\mu/2)$
 \; $C(x)=m+2\cdot R(x)$ 
\newline \phantom{x}\hspace{3ex} return  $c=C(\alpha)$ mod $p$;
\newline \texttt{Decrypt}($c, sk$):
\newline \phantom{x}\hspace{3ex} $m = (c - \lfloor c \cdot B/p \rceil )$ mod $2$;
\phantom{x}\hspace{1ex} return $m$;
\newline \texttt{Add}($c_1$, $c_2$, $pk$):
\newline \phantom{x}\hspace{3ex} $c_3=c_1+c_2$ mod $p$; 
\phantom{x}\hspace{1ex}  return $c_3$;
\newline \texttt{Mult}($c_1$, $c_2$, $pk$):
 \newline \phantom{x}\hspace{3ex}  $c_3=c_1.c_2$ mod $p$; 
\phantom{x}\hspace{1ex}  return $c_3$;

\textbf{Security of the scheme :} The scheme is semantically secure under the assumption that the \textit{Polynomial Coset Problem}(PCP) is hard. One wayness of the encryption is believed to be true under the assumption that the \textit{Bounded distance decoding problem} BDDP is hard while key recovery is assumed to be difficult under the assumption that the \textit{Small Principal Ideal Problem}(SPIP) is hard. Solutions to all of these problems have either a sub-exponential algorithm or in general take exponential time.

@@

\textbf{SWHE to FHE :} We recall that the authors of the scheme have proved that the multiplicative depth $d$ of SWHE scheme verifies: \[d \cdot \mathrm{log}(2) \leq  \mathrm{log}\left( \mathrm{log}\left(\frac{\nu}{2\cdot \sqrt{N}}\right) \right) - \mathrm{log}\left(\mathrm{log}\left( N\mu\right)\right)\]
when $F(x)= x^N+1$. Turning the SWHE to FHE is achieved by recasting Gentry's method for the parameters of this scheme. To achieve this, the authors propose to generate $s_1$ uniformly random integers $B_i$ in $[-p, \ldots, p]$ such that there exists a subset $S$ of $s_2$ elements with $\sum_{j \in S}{B_j}= B$ over integers. By defining $sk_i =1$ iff $i\in S$ and taking $c_i=\texttt{Encrypt}(sk_i,pk)$ we generate a new public key which contains hint of the secret key $sk$. The public key now comprises of \[(p, \alpha, s_1, s_2, \{c_i, B_i\}_{i=1}^{s1})\]
To ensure that the hint does not provide any considerable advantage to an attacker, the underlying \textit{sparse subset-sum problem} SSSP must remain remain hard. If we take $s_1$ to be slightly greater than $p$, then we need to select $s_2$ such that $\sqrt {{s_1}  \choose {s_2} } \leq 2 ^{N/\lambda}$ so as to ensure that SSSP difficulty is at least as difficult as the difficulty of the BDDP underlying the SWHE scheme.

\subsection{Considered Implementations}
We used the open source implementation Scarab library \cite{hcrypt}  version $1.0.0$ of the Smart-Vercauteren scheme which is a part of \textsc{hcryptProject}. The implementation is in beta version. The software requires GMP(we tested with version $5.1.2$) \cite{gmp} : GNU Multiple Precision Arithmetic Library; FLINT version $1.6$ \cite{flint}: Fast Library for Number Theory; MPIR (we tested with $2.6.0$) \cite{mpir} : Multiple Precision Integers and Rationals and MPFR (tested with version $3.1.2$) \cite{mpfr} :  Multiple-Precision Floating-point computations with correct Rounding. Linux is supposed to be the only supported platform. The libraries are implemented in  C and an installation guide is provided with the package.

\autoref{funs} presents the list of functions provided by the implementation with their description. The library provides a file \texttt{test.c} where a user can write his own functions to test different operations. It also provides a header file \texttt{parameters.h} to set the parameters $N, \nu, \mu, s_1$ and $ s_2$ of the cryptosystem. To simplify our computational circuit, we added two basic functions \texttt{or(mpz\_t res, mpz\_t a, mpz\_t b, fhe\_pk\_t pk);} and \texttt{not(mpz\_t res, mpz\_t a, fhe\_pk\_t pk);} which compute the logical \texttt{or}(+) of encrypted bits and the negation ($\neg$) of an encrypted bit respectively.

\begin{table*}
\centering
\caption{Functions available in Scarab}
\begin{tabular}{|l|l||}
  \hline
  \textbf{function prototype} & \textbf{semantics}  \\
  \hline
 \texttt{fhe\_keygen(fhe\_pk\_t pk, fhe\_sk\_t sk);}  & 	Generate a keypair \\
\texttt{fhe\_encrypt(mpz\_t c, fhe\_pk\_t pk, int m);} &	Encrypt a message (0 or 1) \\
\texttt{fhe\_decrypt(mpz\_t c, fhe\_sk\_t sk);} &	Decrypt a cyphertext\\
\texttt{fhe\_recrypt(mpz\_t c, fhe\_pk\_t pk, fhe\_sk\_t sk); }	&Recrypt a cyphertext ( ``refreshing'' it) \\
\texttt{fhe\_add(mpz\_t res, mpz\_t a, mpz\_t b, fhe\_pk\_t pk);} &	Add cypher-texts (= XOR) \\
\texttt{fhe\_mul(mpz\_t res, mpz\_t a, mpz\_t b, fhe\_pk\_t pk);} &	Multiply cypher-texts (= AND) \\
\texttt{fhe\_fulladd(mpz\_t sum, mpz\_t c\_out, mpz\_t a, mpz\_t b,} &	Add with carry in and carry out \\
\phantom{x}\hspace{12ex} \texttt{ mpz\_t c\_in, fhe\_pk\_t pk);} & \\
\texttt{fhe\_halfadd(mpz\_t sum, mpz\_t c\_out, mpz\_t a,} &	Add with carry out \\
\phantom{x}\hspace{12ex} \texttt{ mpz\_t b, fhe\_pk\_t pk);} & \\  
\hline
\end{tabular}
\label{funs}
\end{table*}

We time the various algorithms on two desktop machines. Both these machines were x86-64 platform running Linux kernel $3.2.0$-$45$-generic  and used GCC $4.6.3$ C compiler.  The first machine housed Intel Core i3 M330 (2.13GHz x 4) processor and will be referred in this article as machine M1 while the other used  housed Intel Core i5 M2430 (2.40GHz x 4) processor and will be referred as machine M2. During the experiments these machines were not connected to any network and were running ony the most needed applications.

\subsection{Initial set-up cost}
Here we mesure the performance in terms of run time for KeyGen, Encrypt, Add, Mult, Recrypt, Decrypt.\\
\begin{table}[htb]
  \centering
  \caption{Run time (ms)}
\resizebox{9 cm}{!}{
\begin{tabular}{|c|c|c|c|c|c|c|c|}
  \hline
  \multicolumn{8}{|c|}{$\mu = 4$ , $\eta = 2^{384}$} \\
  \hline
  $N=2^{n}$ & $log_{2}(p)$ & KeyGen & Encrypt & Add & Mult & Recrypt & Decrypt \\
  \hline
  $2^{1}$ & 764 & 99 & 0.057 & 0.0 & 2.372 & 2.371 & 0.0 \\
  $2^{2}$ & 1535 & 221 & 0.072 & 0.0 & 	3.012 & 3.023 & 0.001\\
  $2^{3}$ & 3077 & 1131 & 0.079 & 0.001 & 3.396 & 3.380 & 0.004 \\
  $2^{4}$ & 6155 & 8560 & 0.135 & 0.002 & 5.881 & 5.889 & 0.012\\
  $2^{5}$ & 12329 & 852849 & 0.477 & 0.005 & 20.230 & 20.057 & 0.028\\
  \hline
\end{tabular}
\normalsize
\label{common cost}
}
\end{table}

@@

\textbf{Encryption and Decryption}\\
\section{Circuit Transformation}
As previously discussed, a user sends his program which runs on unencrypted data and his encrypted input to the cloud. The cloud then performs the demanded computation and returns the encrypted result. However, the program in the same form cannot always be used for computations on encrypted data.

Consider a program which involves the following test on two inputs : \texttt{if(a > b) instruction\_1; else instruction\_2;}. As the cloud does not know the result of the test\texttt{ a' > b'} where \texttt{a'} and \texttt{b'} are the encryption of \texttt{a} and \texttt{b}, he does not know which of the instructions he has perform. In fact, he performs both the instructions by transforming the actual program into : \texttt{test(a'> b')*} \texttt{instruction\_1 +}$\neg$\texttt{test(a'> b')*instruction\_2}, where \texttt{test()}  is a function on encrypted inputs \texttt{a'}, \texttt{b'} and returns an encryption of $1$ iff \texttt{a > b}. Hence, if the program used has branching i.e. it involves \texttt{if(cond\_0)} \texttt{\{inst\_0\} else if(cond\_1)\{inst\_1\}} \texttt{ ...else if(cond\_n-2) \{inst\_n-2\}} \texttt{else\{inst\_n-1 \}}; the cloud has to perform all the $n$ instructions i.e. the size of the circuit changes from a constant size to a size linear in the number of tests.

@@

\textbf{Another transformation}: We consider the following program that a cloud receives : \\
 \phantom{x} \texttt{i=0, res=0;}
\newline \phantom{x} \texttt{do\{ }
\newline \phantom{x} \hspace{9ex} \texttt{i++;} 
\newline   \phantom{x} \} \texttt{while(i < n \&\& cond\_i);} 
\newline  \phantom{x} \texttt{res = f(i)};\\
In this case, the cloud needs to transform a loop with conditional stopping into a loop with unconditional branching. We propose the following transformation : \\
 \phantom{x} \texttt{endLoop=0', res=0', i=0;}
\newline \phantom{x} \texttt{do\{ }
\newline \phantom{x} \hspace{6ex} \texttt{endLoop = $\neg$cond\_i;}
\newline \phantom{x} \hspace{6ex} \texttt{res = $\neg$endLoop*f(i)+($\neg$cond\_i)*res;}
\newline \phantom{x} \hspace{6ex} \texttt{i++;} 
\newline   \phantom{x} \} \texttt{while(i < n);}

\textbf{Correctness of the code :} If \texttt{cond\_i} is always \texttt{true}, \texttt{endLoop} is always \texttt{false} and \texttt{res=f(n)}. If at the $t^{th}$ iteration, \texttt{cond\_i} is \texttt{false}, \texttt{endLoop} becomes \texttt{true} and hence \texttt{res} contains its previous value $f(t-1)$.

@@

\subsection{ Didactic example :  Min-Max}
\label{sec:ex}
Suppose that a user wants to find the maximum and the minimum of two $n$-bit sequences $a = (a_0,a_1,\ldots, a_{n-1})$, $b=(b_0,b_1,\ldots, b_{n-1})$ where $a_0$ is the highest significant bit. He provides the following algorithm to the cloud :
\newline \phantom{x} \texttt{aIsGreater=0, i=0;}
\newline \phantom{x} \texttt{do\{ }
\newline \phantom{x} \hspace{6ex} \texttt{ if(a[i]==1 \&\& b[i]==0) 
\newline \phantom{x} \hspace{14ex}aIsGreater = 1;}
\newline \phantom{x} \hspace{7ex} \texttt{i++;} 
\newline   \phantom{x} \} \texttt{while(i < n);}

@@

 \caption{Min-Max on cipher-text \label{Code:algo}}

@@

  Consider $x_i = a_i'b_i'+\neg(a_i')\neg(b_i') $. $x_i$ is $1$ iff $a_i'$ and $b_i'$ are equal. We note that, $bitEqual$ is the product of $x_i$ and is $1$ at the $t^{th}$ iteration iff $a_i'=b_i'$ for all $i < t$. The term $\neg(b_i')a_i'$ is $1$ iff $a_i'=1$ and $b_i'=0$. Hence at the $t^{th}$ iteration, the algorithm using $bitEqual$ tests if the previous bits were equal. Hence, $aIsGreater$ is $1$ iff at any stage  $\neg(b_t')a_t'=1$ and the previous bits were equal.

Max and Min regenerate $a_i$ and $b_i$ and hence contain the correct result.  
\end{proof}
\textbf{Theoretical Cost :} Algorithm \autoref{Code:algo} uses $5n$ \texttt{AND} gates, $2n$ \texttt{OR} gates and $2n$ \texttt{NOT} gates to find the larger of the two bit sequences and then to regenerate the maximum and the minimum $4n$ \texttt{AND} gates and $2n$ \texttt{NOT} and \texttt{OR} gates. We note that reconstruction of Max and  Min is not required when working on non-encrypted data. Hence, this is an additional cost to pay to use the transformed program operating on encrypted data.

\textbf{Experimental Analysis :} We measure the run time from $n=1$ to $32$ bits, we find a linear cost which confirms the $\theta(n)$ complexity. During these measures the result of Min-Max is correct.
\begin{comment} 
\begin{figure}[!h] %on ouvre l'environnement figure
\centering
\includegraphics[width=5cm]{f2.eps} 
\caption{Run time(ms)} 
\label{image_f2} %l'étiquette pour faire référence à cette image
\end{figure}
\end{comment}
The cost of this algorithm on cipher-text is linear with bit number, like  the algorithm on plain-text, the extra cost in this case is encryption-decryption cost and especially the code generation which is very unpredictable and take often a long time in the SV scheme.

@@

We propose a set of benchmark functions to evaluate the implementations. These functions can be categorized into three classes : functions operating on bits, functions operating on block of bits and functions where branching is usually required i.e \texttt{if-then-else} condition is evaluated.
